An individual differences perspective on the development of pragmatic reasoning

### 2022-02-19 paper/manuscript.tex
add age models

---
title             : "An individual differences perspective on the development of pragmatic reasoning"
shorttitle        : "Individual differences in pragmatic reasoning"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany"
    email         : "manuel_bohn@eva.mpg.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Michael Henry Tessler"
    affiliation   : "2,3"
    role:
      - Analysis
      - Writing - Review & Editing
  - name          : "Clara Kordt"
    affiliation   : "4"
    role:
      - Data collection
  - name          : "Tom Hausmann"
    affiliation   : "5"
    role:
      - Data collection
  - name          : "Michael C. Frank"
    affiliation   : "6"
    role:
      - Conceptualization
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
  - id            : "2"
    institution   : "DeepMind, London, UK"
  - id            : "3"
    institution   : "Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology"
  - id            : "4"
    institution   : "Martin Luther University Halle-Wittenberg"
  - id            : "5"
    institution   : "Brandenburg Medical School Theodor Fontane"
  - id            : "6"
    institution   : "Department of Psychology, Stanford University"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

  M. Bohn received funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Sklodowska-Curie grant agreement no. 749229. M. H. Tessler was funded by the National Science Foundation SBE Postdoctoral Research Fellowship Grant No. 1911790. M. C. Frank was supported by a Jacobs Foundation Advanced Research Fellowship and the Zhou Fund for Language and Cognition. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : "../../../References/library.bib"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)
library(knitr)
library(brms)

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r}
data <- bind_rows(
  read_csv("../data/data_r1.csv") %>%filter(task != "training") %>%mutate(study = "study1"),
  read_csv("../data/data_r2.csv") %>%filter(task != "training")%>%mutate(study = "study2"),
  read_csv("../data/data_r3.csv") %>%filter(task != "training")%>%mutate(study = "study3")
  ) %>%
  mutate(study = factor(study),
         id = paste(study, id, sep = "_"), 
         test_day = paste0("day", test_day))

```
# Introduction

# Facets of pragmatic reasoning

# The challenges of studying individual differences

# The current study

# Study 1

Methods and sample size were pre-registered at https://osf.io/6a723. All analysis scripts and data files can be found in the following repository: https://github.com/manuelbohn/pragBat. The same repository also contains the code to run the experiments.

## Participants
```{r}
dem <- data%>%
  group_by(id)%>%
  summarise(study = unique(study),
            gender = unique(gender),
            age = unique(age),
            testdays = length(unique(test_day)))%>%
  group_by(study)%>%
  summarise(n = length(id),
            reli_n = sum(testdays == 2),
            female = sum(gender == "f"),
            mage = mean(age),
            lrange = range(age)[1],
            urange = range(age)[2])
```

For Study 1, we collected data from `r dem%>%filter(study == "study1")%>%pull(n)` children ($m_{age}$ = `r dem%>%filter(study == "study1")%>%pull(mage)`, range$_{age}$: `r dem%>%filter(study == "study1")%>%pull(lrange)` - `r dem%>%filter(study == "study1")%>%pull(urange)`, `r dem%>%filter(study == "study1")%>%pull(female)` girls) of which `r dem%>%filter(study == "study1")%>%pull(reli_n)` were tested twice. For most children, the two test sessions were two days apart; the longest time difference was six days. Children came from an ethnically homogeneous, mid-size German city (~550,000 inhabitants, median income €1,974 per month as of 2020); were mostly monolingual and had mixed socioeconomic backgrounds. The study was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology. Data was collected between November 2019 and January 2020.

## Material and Methods

The study was presented as an interactive picture book on a tablet computer. The tasks were programmed in `HTML/JavaScript` and run in a web browser. Pre-recorded sound files were used to address the child (one native German speaker per animal). Children responded by touching objects on the screen. Children were tested in a quiet room in their daycare or in a separate room in a child laboratory. An experimenter guided the child through the study, selecting the different tasks and advancing within each task. In the beginning of the study, children completed a touch training to familiarize themselves with selecting objects. After a short introduction to the different animal characters, children completed the following six tasks. Figure \@ref(fig:fig1) shows screenshots for each task and the order in which they were presented.

### Training

An animal was standing on a pile between two tables. On each table, a familiar object was located. The animal asked the child to give them one of the objects (e.g., "Can you give me the car"). the objects were chosen so that children of the youngest age group would easily understand them (car and ball). This procedure familiarized the child with the general logic of the animals making requests and the child touching objects. There were two training trials.

### Mutual exclusivity

This task was directly taken from @bohn2021young. The task layout and the procedure was the same as in the training. In each trial, one object was a novel object (drawn for the purpose of this study) while the other one was likely to be familiar to children. Both object types changed from trial to trial. Following @bohn2021young, the familiar objects varied in terms of the likelihood that they would be familiar to children in the age range (carrot, duck, eggplant, garlic, horseshoe). For example, we assumed that most 3-year-olds would recognize a carrot, whereas fewer children would recognize a horseshoe. The animal always used a novel non-word (e.g., gepsa) in their request. We reasoned that children would identify the novel object as the referent of the novel word because they assumed the animal would have used the familiar word if they wanted to request the familiar object. Children's response was thus coded as correct if they selected the novel object. There were five trials, with the side on which the novel object appeared pseudo-randomized.

### Informativeness inference

The task was directly taken from @bohn_tessler_merrick_frank_2019. Th animal was standing between two trees with objects hanging in them. In one tree, there were two objects (type A and B) and in the other tree there was only one (type B). The animal turned to the tree with the two objects and labelled one of the objects. It was unclear from the animal's utterance, which of the two objects they were referring to. We assumed that children would map the novel word onto the object of type A because they expected the animal to turn to the tree with only the object of type B if their intention was to provide a label for an object of type B. Next, the trees were replaced by new ones, one of which carried an object of type A and the other of type B. The animal then said that one of the trees had the same object as they labelled previously (using the same label) and asked the child to touch the tree. We coded as correct if the child selected the tree with the object of type A. The first two trials were training trials, in which there was only one object in each tree. There were five test trials. The location of the tree with the two objects in the beginning of each trial was pseudo-randomized and so was the location of the objects when the new trees appeared.

### Speaker preference

This task was also taken from @bohn_tessler_merrick_frank_2019. The animal was standing between the two tables, each of which had a novel object (drawn for the purpose of the study) on it. The animal turned to one table, pointed at the object and said that they very much liked this object (using a pronoun instead of a label). Next, the animal turned to the other table and said that they really did not like the object (again, using a pronoun and no label). Then the animal turned towards the participant and used a novel label to request an object in an excited tone. We assumed that children would track the animal's preference and identify the previously liked object as the referent. Thus, we coded as correct if the child selected the object the animal expressed preference for. There were five test trials. The location of the preferred object as well as whether the animal first expressed liking or disliking was pseudo-randomized across trials

### Discourse novelty

This task was taken from @bohn2021young. Once again, the animal was standing between the two tables. One table was empty whereas there was a novel object on the other table. The animal turned towards the empty table and commented on its emptiness. Next, the animal turned to the other table and commented (in a neutral tone) on the presence of the object (not using a label). The animal then briefly disappeared. In the absence of the animal a second novel object appeared on the previously empty table. Then the animal returned and, facing the participant, asked for an object in an excited tone. We assumed that children would track which object was new to the ongoing interaction and identify the object that was new in context as the referent. We coded as correct when children selected the object that appeared later. There were five test trials. The location of the empty table and whether the animal first commented on the presence or absence of an object was pseudo-randomized across trials

### Card sorting

This task was modeled after @zelazo2006dimensional. The child saw to cards, a blue rabbit on the left and a red boat on the right. The experimenter introduced the child to the color game they would be playing next. In this game, all blue cards (irrespective of object depicted) would go to the left card and all red cards to the right. Next, a third card appeared in the middle of the screen (red rabbit or blue boat) and the experimenter demonstrated the color sorting by moving the card to the one with the same color. After a second demonstration trial, the child started to do the color sorting by themselves. After six trials, the experimenter said that they were now going to play a different game, the shape game, according to which all rabbits would go to the card with the rabbit (left) and all boats to the card with the boat (right). The experimenter repeated these instructions once and without any demonstration the child continued with the sorting according to the new rule. There were six test trials. The shape on the card was pseudo-randomized across trials. We only coded the trials after the rule change and coded as correct when the child sorted according to shape.

@@

(ref:figlab1) Overview of the tasks used in Study 1 to 3. Pictures show screenshots from each task. The vertical order corresponds to the order of presentation in each study. The colors group the tasks along the (Assumed) cognitive processes involved. Blue: utterance-based inferences. Red common ground/discourse-based inferences. Green: Executive functions. Green: Analogical reasoning.

```{r fig1, include = T, fig.align = "center", fig.cap = "(ref:figlab1)", out.width="100%"}
knitr::include_graphics("./figures/figure1.png")
```

## Analysis

@@

All analysis were run in `R` [@R-base] version 4.1.2. Regression models were fit as Bayesian generalized linear mixed models (GLMM) using the function `brm` from the package `brms` [@burkner2017brms]. We used default priors for all analysis.

To estimate developmental effects in each task, we fit a GLMM predicting correct responses (0/1) by age (centered at the mean) and trial number (also centered). The model included random intercepts for each participant and random slopes for trial within participants (model notation in `R`: `correct ~ age + trial + (trial|id)` ). For each task, we inspected and visualized the posterior distribution (mean and 95% Credible Interval (CrI)) for the age estimate.

We assessed re-test reliability in two ways. First, for each task we computed the proportion of correct trials for each individual in the two test sessions and then used  Pearson correlations to quantify re-test reliability. Second, we used a GLMM based approach suggested by @rouder2019psychometrics. Here, a GLMM was fitted to the trial-by-trial data for each task with a fixed effect of age, a random intercept for each participant and a random slope for test day (`correct ~ age + (0+test_day|id)`)^[The notation `0+test_day` yields a separate intercept estimate for each test day and subject instead of an intercept estimate for day 1 and a slope for the difference between day 1 and day 2. As a consequence, the model estimates the correlation between the two test days instead of a correlation between an intercept and the slope for test day.]. The model yields a participant specific estimate for each test day and also estimates the correlation between the two. This correlation can be interpreted as the re-test reliability. This approach has several advantages. First, it uses the trial-by-trial data and avoids information loss that comes with data aggregation. Second, it uses hierarchical shrinkage to obtain better participant specific estimates. Finally, it allows us to get an age-independent estimate for reliability. One worry when assessing re-test reliability in developmental studies is that re-test correlations can be high because of domain general cognitive gains and not because of task-specific individual differences. By including age as a fixed effect in the model, the estimates for each participant are independent of age and so is the correlation between estimates for the two test days -- the re-test reliability.

@@

## Results

We found developmental effects in most of the tasks. Figure \@ref(fig:fig2) shows the data and visualizes the developmental trajectories based on the model. Figure \@ref(fig:fig3) shows the model estimates for age. In the mutual exclusivity task, performance was reliably above chance level and increased with age. For informative inference, the pattern was quite different: Performance was at chance level with only minor developmental gains. In the speaker preference task, performance was again clearly above chance with developmental gains resulting in a ceiling effect for older children. In the discourse novelty task, performance was also above chance with no clear developmental effects. The card sorting task showed the strongest developmental effects with younger children performing largely below chance and older children performing above chance.

Re-test reliability was high for most tasks (see Figure \@ref(fig:fig2)). Raw correlations between the two test sessions was above .7 for mutual exclusivity, speaker preference and discourse novelty. With .62 it was slightly lower for card sorting. The model based -- age independent -- reliability estimates yielded similar results suggesting that the tasks did capture task specific individual differences. A notable exception was the informativeness inference task, which was not reliable according to any of the methods of computing re-test correlations. We suspect the overall low variation in performance to be responsible for this.

```{r, echo = F}
cors <- read_csv("../saves/by_task_correlations.csv")
```

Most correlations between the tasks were low and ranged between *r* = -0.2 and 0.2 (see Figure \@ref(fig:fig2)). A notable exception was the correlation between mutual exclusivity and card sorting with *r* = `r cors%>%filter(study == "study1", task1 == "mutual_exclusivity", task2 == "card_sorting")%>%pull(rho)` (95% CI[`r cors%>%filter(study == "study1", task1 == "mutual_exclusivity", task2 == "card_sorting")%>%pull(lci)` - `r cors%>%filter(study == "study1", task1 == "mutual_exclusivity", task2 == "card_sorting")%>%pull(uci)`]).

(ref:figlab2) Results by task for studies 1 to 3. Each panel shows the results for one task. Regression lines show the predicted developmental trajectories (with 95% CrI) based on by-task GLMMs, with the line type indicating the study. Colored points show age group means (with 95% CI based on non-parametric bootstrap) with the different shapes corresponding to the different studies. Light shapes show the mean performance for each subject by study. Dotted line shows level of performance expected by chance.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{An individual differences perspective on the development of pragmatic reasoning}
\author{Manuel Bohn\textsuperscript{1}, Michael Henry Tessler\textsuperscript{2,3}, Clara Kordt\textsuperscript{4}, Tom Hausmann\textsuperscript{5}, \& Michael C. Frank\textsuperscript{6}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An individual differences perspective on the development of pragmatic reasoning},
  pdfauthor={Manuel Bohn1, Michael Henry Tessler2,3, Clara Kordt4, Tom Hausmann5, \& Michael C. Frank6},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Individual differences in pragmatic reasoning}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

M. Bohn received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Sklodowska-Curie grant agreement no. 749229. M. H. Tessler was funded by the National Science Foundation SBE Postdoctoral Research Fellowship Grant No.~1911790. M. C. Frank was supported by a Jacobs Foundation Advanced Research Fellowship and the Zhou Fund for Language and Cognition. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

The authors made the following contributions. Manuel Bohn: Conceptualization, Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Henry Tessler: Analysis, Writing - Review \& Editing; Clara Kordt: Data collection; Tom Hausmann: Data collection; Michael C. Frank: Conceptualization, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn, Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany. E-mail: \href{mailto:manuel_bohn@eva.mpg.de}{\nolinkurl{manuel\_bohn@eva.mpg.de}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany\\\textsuperscript{2} DeepMind, London, UK\\\textsuperscript{3} Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology\\\textsuperscript{4} Martin Luther University Halle-Wittenberg\\\textsuperscript{5} Brandenburg Medical School Theodor Fontane\\\textsuperscript{6} Department of Psychology, Stanford University}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{facets-of-pragmatic-reasoning}{%
\section{Facets of pragmatic reasoning}\label{facets-of-pragmatic-reasoning}}

\hypertarget{the-challenges-of-studying-individual-differences}{%
\section{The challenges of studying individual differences}\label{the-challenges-of-studying-individual-differences}}

\hypertarget{the-current-study}{%
\section{The current study}\label{the-current-study}}

\hypertarget{study-1}{%
\section{Study 1}\label{study-1}}

Methods and sample size were pre-registered at \url{https://osf.io/6a723}. All analysis scripts and data files can be found in the following repository: \url{https://github.com/manuelbohn/pragBat}. The same repository also contains the code to run the experiments.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

For Study 1, we collected data from 48 children (\(m_{age}\) = 3.99, range\(_{age}\): 3.10 - 4.99, 23 girls) of which 41 were tested twice. For most children, the two test sessions were two days apart; the longest time difference was six days. Children came from an ethnically homogeneous, mid-size German city (\textasciitilde550,000 inhabitants, median income €1,974 per month as of 2020); were mostly monolingual and had mixed socioeconomic backgrounds. The study was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology. Data was collected between November 2019 and January 2020.

\hypertarget{material-and-methods}{%
\subsection{Material and Methods}\label{material-and-methods}}

The study was presented as an interactive picture book on a tablet computer. The tasks were programmed in \texttt{HTML/JavaScript} and run in a web browser. Pre-recorded sound files were used to address the child (one native German speaker per animal). Children responded by touching objects on the screen. Children were tested in a quiet room in their daycare or in a separate room in a child laboratory. An experimenter guided the child through the study, selecting the different tasks and advancing within each task. In the beginning of the study, children completed a touch training to familiarize themselves with selecting objects. After a short introduction to the different animal characters, children completed the following six tasks. Figure \ref{fig:fig1} shows screenshots for each task and the order in which they were presented.

\hypertarget{training}{%
\subsubsection{Training}\label{training}}

An animal was standing on a pile between two tables. On each table, a familiar object was located. The animal asked the child to give them one of the objects (e.g., ``Can you give me the car''). the objects were chosen so that children of the youngest age group would easily understand them (car and ball). This procedure familiarized the child with the general logic of the animals making requests and the child touching objects. There were two training trials.

\hypertarget{mutual-exclusivity}{%
\subsubsection{Mutual exclusivity}\label{mutual-exclusivity}}

This task was directly taken from Bohn, Tessler, Merrick, and Frank (2021). The task layout and the procedure was the same as in the training. In each trial, one object was a novel object (drawn for the purpose of this study) while the other one was likely to be familiar to children. Both object types changed from trial to trial. Following Bohn et al. (2021), the familiar objects varied in terms of the likelihood that they would be familiar to children in the age range (carrot, duck, eggplant, garlic, horseshoe). For example, we assumed that most 3-year-olds would recognize a carrot, whereas fewer children would recognize a horseshoe. The animal always used a novel non-word (e.g., gepsa) in their request. We reasoned that children would identify the novel object as the referent of the novel word because they assumed the animal would have used the familiar word if they wanted to request the familiar object. Children's response was thus coded as correct if they selected the novel object. There were five trials, with the side on which the novel object appeared pseudo-randomized.

\hypertarget{informativeness-inference}{%
\subsubsection{Informativeness inference}\label{informativeness-inference}}

The task was directly taken from Bohn, Tessler, Merrick, and Frank (2022). Th animal was standing between two trees with objects hanging in them. In one tree, there were two objects (type A and B) and in the other tree there was only one (type B). The animal turned to the tree with the two objects and labelled one of the objects. It was unclear from the animal's utterance, which of the two objects they were referring to. We assumed that children would map the novel word onto the object of type A because they expected the animal to turn to the tree with only the object of type B if their intention was to provide a label for an object of type B. Next, the trees were replaced by new ones, one of which carried an object of type A and the other of type B. The animal then said that one of the trees had the same object as they labelled previously (using the same label) and asked the child to touch the tree. We coded as correct if the child selected the tree with the object of type A. The first two trials were training trials, in which there was only one object in each tree. There were five test trials. The location of the tree with the two objects in the beginning of each trial was pseudo-randomized and so was the location of the objects when the new trees appeared.

\hypertarget{speaker-preference}{%
\subsubsection{Speaker preference}\label{speaker-preference}}

This task was also taken from Bohn et al. (2022). The animal was standing between the two tables, each of which had a novel object (drawn for the purpose of the study) on it. The animal turned to one table, pointed at the object and said that they very much liked this object (using a pronoun instead of a label). Next, the animal turned to the other table and said that they really did not like the object (again, using a pronoun and no label). Then the animal turned towards the participant and used a novel label to request an object in an excited tone. We assumed that children would track the animal's preference and identify the previously liked object as the referent. Thus, we coded as correct if the child selected the object the animal expressed preference for. There were five test trials. The location of the preferred object as well as whether the animal first expressed liking or disliking was pseudo-randomized across trials

\hypertarget{discourse-novelty}{%
\subsubsection{Discourse novelty}\label{discourse-novelty}}

This task was taken from Bohn et al. (2021). Once again, the animal was standing between the two tables. One table was empty whereas there was a novel object on the other table. The animal turned towards the empty table and commented on its emptiness. Next, the animal turned to the other table and commented (in a neutral tone) on the presence of the object (not using a label). The animal then briefly disappeared. In the absence of the animal a second novel object appeared on the previously empty table. Then the animal returned and, facing the participant, asked for an object in an excited tone. We assumed that children would track which object was new to the ongoing interaction and identify the object that was new in context as the referent. We coded as correct when children selected the object that appeared later. There were five test trials. The location of the empty table and whether the animal first commented on the presence or absence of an object was pseudo-randomized across trials

\hypertarget{card-sorting}{%
\subsubsection{Card sorting}\label{card-sorting}}

This task was modeled after Zelazo (2006). The child saw to cards, a blue rabbit on the left and a red boat on the right. The experimenter introduced the child to the color game they would be playing next. In this game, all blue cards (irrespective of object depicted) would go to the left card and all red cards to the right. Next, a third card appeared in the middle of the screen (red rabbit or blue boat) and the experimenter demonstrated the color sorting by moving the card to the one with the same color. After a second demonstration trial, the child started to do the color sorting by themselves. After six trials, the experimenter said that they were now going to play a different game, the shape game, according to which all rabbits would go to the card with the rabbit (left) and all boats to the card with the boat (right). The experimenter repeated these instructions once and without any demonstration the child continued with the sorting according to the new rule. There were six test trials. The shape on the card was pseudo-randomized across trials. We only coded the trials after the rule change and coded as correct when the child sorted according to shape.

@@

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure1}

}

\caption{Overview of the tasks used in Study 1 to 3. Pictures show screenshots from each task. The vertical order corresponds to the order of presentation in each study. The colors group the tasks along the (Assumed) cognitive processes involved. Blue: utterance-based inferences. Red common ground/discourse-based inferences. Green: Executive functions. Green: Analogical reasoning.}\label{fig:fig1}
\end{figure}

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

@@

All analysis were run in \texttt{R} (R Core Team, 2018) version 4.1.2. Regression models were fit as Bayesian generalized linear mixed models (GLMM) using the function \texttt{brm} from the package \texttt{brms} (Bürkner, 2017). We used default priors for all analysis.

To estimate developmental effects in each task, we fit a GLMM predicting correct responses (0/1) by age (centered at the mean) and trial number (also centered). The model included random intercepts for each participant and random slopes for trial within participants (model notation in \texttt{R}: \texttt{correct\ \textasciitilde{}\ age\ +\ trial\ +\ (trial\textbar{}id)} ). For each task, we inspected and visualized the posterior distribution (mean and 95\% Credible Interval (CrI)) for the age estimate.

We assessed re-test reliability in two ways. First, for each task we computed the proportion of correct trials for each individual in the two test sessions and then used Pearson correlations to quantify re-test reliability. Second, we used a GLMM based approach suggested by Rouder and Haaf (2019). Here, a GLMM was fitted to the trial-by-trial data for each task with a fixed effect of age, a random intercept for each participant and a random slope for test day (\texttt{correct\ \textasciitilde{}\ age\ +\ (0+test\_day\textbar{}id)})\footnote{The notation \texttt{0+test\_day} yields a separate intercept estimate for each test day and subject instead of an intercept estimate for day 1 and a slope for the difference between day 1 and day 2. As a consequence, the model estimates the correlation between the two test days instead of a correlation between an intercept and the slope for test day.}. The model yields a participant specific estimate for each test day and also estimates the correlation between the two. This correlation can be interpreted as the re-test reliability. This approach has several advantages. First, it uses the trial-by-trial data and avoids information loss that comes with data aggregation. Second, it uses hierarchical shrinkage to obtain better participant specific estimates. Finally, it allows us to get an age-independent estimate for reliability. One worry when assessing re-test reliability in developmental studies is that re-test correlations can be high because of domain general cognitive gains and not because of task-specific individual differences. By including age as a fixed effect in the model, the estimates for each participant are independent of age and so is the correlation between estimates for the two test days -- the re-test reliability.

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

We found developmental effects in most of the tasks. Figure \ref{fig:fig2} shows the data and visualizes the developmental trajectories based on the model. Figure \ref{fig:fig3} shows the model estimates for age. In the mutual exclusivity task, performance was reliably above chance level and increased with age. For informative inference, the pattern was quite different: Performance was at chance level with only minor developmental gains. In the speaker preference task, performance was again clearly above chance with developmental gains resulting in a ceiling effect for older children. In the discourse novelty task, performance was also above chance with no clear developmental effects. The card sorting task showed the strongest developmental effects with younger children performing largely below chance and older children performing above chance.

Re-test reliability was high for most tasks (see Figure \ref{fig:fig2}). Raw correlations between the two test sessions was above .7 for mutual exclusivity, speaker preference and discourse novelty. With .62 it was slightly lower for card sorting. The model based -- age independent -- reliability estimates yielded similar results suggesting that the tasks did capture task specific individual differences. A notable exception was the informativeness inference task, which was not reliable according to any of the methods of computing re-test correlations. We suspect the overall low variation in performance to be responsible for this.

Most correlations between the tasks were low and ranged between \emph{r} = -0.2 and 0.2 (see Figure \ref{fig:fig2}). A notable exception was the correlation between mutual exclusivity and card sorting with \emph{r} = 0.31 (95\% CI{[}0.03 - 0.55{]}.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure2}

}

\caption{Results by task for studies 1 to 3. Each panel shows the results for one task. Regression lines show the predicted developmental trajectories (with 95\% CrI) based on by-task GLMMs, with the line type indicating the study. Colored points show age group means (with 95\% CI based on non-parametric bootstrap) with the different shapes corresponding to the different studies. Light shapes show the mean performance for each subject by study. Dotted line shows level of performance expected by chance.}\label{fig:fig2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{./figures/figure3}

}

\caption{Model estimates (with 95\% CrI) for age based on GLMMs for each task and study.}\label{fig:fig3}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure4}

}

\caption{xxx.}\label{fig:fig4}
\end{figure}
